### 2019-07-03 2-Chapters/2-Chapter/chapter2.tex
Done for 2day

\paragraph{A note about the implementation of \klUCB{} algorithms.}
%
Implementing this kind of policy requires to implement two things.
The first one is usually easy: one has to implement $\kl(x,y)$ the Kullback-Leibler divergence of the considered one-dimensional family,
for instance $\kl$ for Bernoulli distributions has a closed form and is easy to compute.
The second one is less obvious: one has to solve the optimisation problem in \eqref{eq:2:klUCB_index}.
Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\kl(x,\bullet)$ function is usually convex (for any $x$), two practical and mathematically correct solutions are Newton's method or a bisection search.
The search space for $q$ can be restricted to $[ X_k(t) / N_k(t), 1 ]$ instead of $[0,1]$,
and for simplicity and efficiency, in this case a naive bisection search works well.
% and Newton's method guarantees that a small number of steps yield a given precision level $\varepsilon$.
For instance, a precision level of $\varepsilon=10^{-8}$ is enough for numerical simulations, and typically the bisection search terminates in between $5$ to $20$ steps.

If we perform numerical simulations with \klUCB{} algorithms, we are interested in having efficient implementations, and ideally \klUCB{} should not be significantly slower than \UCB. The bottleneck of the performance is the computations of the $\KL$ functions and the optimization problem,
and we note that our library SMPyBandits implements the two parts in both a naive Python implementation (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback.py}{\texttt{kullback.py}}) and in a Python module written in a
C file\footnote{~We also wrote an intermediate implementation that uses Cython \cite{cython}, to have a file with a syntax closer to Python, while gaining from almost the same performance speed-up as with the optimized C version, see file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback_cython.pyx}{\texttt{kullback\_cython.pyx}}.
    These implementations are also published an independent Python and Julia packages, on \href{https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}{\texttt{GitHub.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}} (also on Pypi),
    and on \href{https://github.com/Naereen/KullbackLeibler.jl}{\texttt{GitHub.com/Naereen/KullbackLeibler.jl}}.
    % and installable with \texttt{Pkg.add("KullbackLeibler")}.
},
in an optimized way (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/C/kullback_py3.c}{\texttt{kullback\_py3.c}}).
For the supported families of distributions (including Bernoulli, Gaussian, exponential and Poisson), we wrote both a naive Python and an optimized CPython version of their Kullback-Leibler divergence (\ie, $\KL(x,y)$) and a bisection search for the optimization problem for the \klUCB{} index (\ie, $\sup_q \KL(x,q) \leq z$).

\paragraph{Variants on $\UCB_1$.}
%
Many variants of the $\UCB_1$ algorithms have been studied in the multi-armed bandits literature.
For instance, \UCB-H replaces the $\log(t)$ by a $\log(T)$ (as well as \klUCB-H), and it actually corresponds to the first algorithm analyzed in \cite{Auer02}.
Using an estimator for the variance of the arms' samples gives \UCB-V, which was proven to be efficient against Gaussian distributions with unknown variance.
Other variants include \UCB$^+$, $\UCB^{\dagger}$ from \cite{Lattimore2018refining}, {\UCB}oost, and many more.
%
All these variants are usually comparable to the original $\UCB_1$ algorithm, in terms of time and memory complexity, and they achieve improved regret upper bounds in the same setting or some extensions of the initial setting.

@@

\begin{table}[!h]
    \begin{small}  % WARNING
        \centering
        \begin{tabular}{c|cc|cc}
        \textbf{Algorithms}
        & Rewards
        & Regret for $T=100$
        & Rewards
        & Regret for $T=10000$ \\
        % & $\sum\limits_{t=1}^{T=100} r(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r(t)$ & $R_{T=10000}^{\cA}$ \\
        % & $\sum\limits_{t=1}^{T=100} r(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r(t)$ & $R_{T=10000}^{\cA}$ \\
            \hline
            Pure exploration
                & $46$ & $20.9$
                & $5171$ & $1886.4$ \\
            \textbf{Pure exploitation}
                & $\mathbf{65}$ & $\mathbf{1.15}$
                & $\mathbf{6997}$ & $\mathbf{0.95}$ \\
            \hline
            $\varepsilon$-greedy
                & $50$ & $20.9$
                & $6657$ & $382.2$ \\
            Explore-then-Commit
                & $49$ & $15$
                & $5766$ & $1199.8$ \\
            \hline
            $\UCB_1$
                & $63$ & $10.3$
                & $6920$ & $78.4$ \\
            \klUCB{}
                & $56$ & $12.4$
                & $6922$ & $70.2$ \\
            \hline
            Thompson sampling
                & $51$ & $12.2$
                & $6949$ & $44.2$ \\
        \end{tabular}
        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{for only one run of the simulation}, for different algorithms.}
        \label{table:2:oneShotResult}
    \end{small}  % WARNING
\end{table}

@@

\begin{table}[ht]
    \begin{small}  % WARNING
        \centering
        \begin{tabular}{c|cc|cc}
        \textbf{Algorithms}
        & Rewards
        & Regret for $T=100$
        & Rewards
        & Regret for $T=10000$ \\
        % & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=100} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=100}$ & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=10000} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=10000}$ \\
            \hline
            Pure exploration
                & $51$ & $19$
                & $5099$ & $1900$ \\
            Pure exploitation
                & $62$ & $8.3$
                & $6349$ & $653$ \\
            \hline
            $\varepsilon$-greedy
                & $51$ & $18.7$
                & $6460$ & $542$ \\
            Explore-then-Commit
                & $57$ & $12.9$
                & $5795$ & $1207$ \\
            \hline
            $\UCB_1$
                & $\mathbf{60}$ & $\mathbf{9.8}$
                & $6921$ & $79$ \\
            \klUCB{}
                & $61$ & $9.3$
                & $6927$ & $72$ \\
            \hline
            Thompson sampling
                & $\mathbf{60}$ & $\mathbf{9.6}$
                & $\mathbf{6951}$ & $\mathbf{49}$ \\
        \end{tabular}
        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{averaged over $N=1000$ independent simulations} ($j=1,\dots,N$), for different algorithms.}
        \label{table:2:meanResults}
    \end{small}  % WARNING
\end{table}

@@

\begin{figure}[h!]  % [htbp]
	% \centering
    \includegraphics[width=1.14\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
	\caption{Average of the cumulated rewards, as function of $t$, for $T=10000$ and $N=1000$.}
	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
\end{figure}

\begin{figure}[h!]  % [htbp]
	% \centering
	\includegraphics[width=1.14\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
\end{figure}

@@

We presented the simplest MAB model studied in this chapter, by focussing on one agent playing a stationary game.
Both hypotheses are removed or weaken in the rest of this manuscript,
% first in Chapter~\ref{chapter:4} where we consider many XXX
first by considering players facing a stationary problem in a decentralized way in Chapter~\ref{chapter:5},
and then by considering a single player facing a non-stationary or a piece-wise stationary problem in Chapter~\ref{chapter:6}.
%
For both directions, we present in the two final chapters natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results for the two problems
of stationary multi-players and piece-wise stationary bandits.

@@

\paragraph{Possible future works.}
%
We focused in this thesis on finite-arms and one-dimensional bandit problems,
and thus two possible directions of future works could be to extend our works
to MAB models with either multidimensional rewards, like contextual bandits, or infinite arms, like Lipschitz bandits.

>>>>

\textbf{A note about the implementation of \klUCB{} algorithms.}
%
Implementing this kind of policy requires to implement two things.
The first one is usually easy: one has to implement $\kl(x,y)$ the Kullback-Leibler divergence of the considered one-dimensional family,
for instance $\kl$ for Bernoulli distributions has a closed form and is easy to compute.
The second one is less obvious: one has to solve the optimisation problem in \eqref{eq:2:klUCB_index}.
Any iterative algorithm for numerical optimization of one-dimensional functions can be used (\eg, see \cite{BoydVanderberghe04} for a survey), but as the $\kl(x,\bullet)$ function is usually convex (for any $x$), two practical and mathematically correct solutions are Newton's method or a bisection search.
The search space for $q$ can be restricted to $[ X_k(t) / N_k(t), 1 ]$ instead of $[0,1]$,
and for simplicity and efficiency, in this case a naive bisection search works well.
% and Newton's method guarantees that a small number of steps yield a given precision level $\varepsilon$.
For instance, a precision level of $\varepsilon=10^{-8}$ is enough for numerical simulations, and typically the bisection search terminates in between $5$ to $20$ steps.

If we perform numerical simulations with \klUCB{} algorithms, we are interested in having efficient implementations, and ideally \klUCB{} should not be significantly slower than \UCB. The bottleneck of the performance is the computations of the $\KL$ functions and the optimization problem,
and we note that our library SMPyBandits implements the two parts in both a naive Python implementation (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback.py}{\texttt{kullback.py}}) and in a Python module written in a
C file\footnote{~We also wrote an intermediate implementation that uses Cython \cite{cython}, to have a file with a syntax closer to Python, while gaining from almost the same performance speed-up as with the optimized C version, see file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/kullback_cython.pyx}{\texttt{kullback\_cython.pyx}}.
    These implementations are also published an independent Python package, on \href{https://github.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}{\texttt{GitHub.com/Naereen/Kullback-Leibler-divergences-and-kl-UCB-indexes}} (also on Pypi),
    as well as a Julia package, on \href{https://github.com/Naereen/KullbackLeibler.jl}{\texttt{GitHub.com/Naereen/KullbackLeibler.jl}}.
    % and installable with \texttt{Pkg.add("KullbackLeibler")}.
},
in an optimized way (file \href{https://github.com/SMPyBandits/SMPyBandits/blob/master/SMPyBandits/Policies/C/kullback_py3.c}{\texttt{kullback\_py3.c}}).
For the supported families of distributions (including Bernoulli, Gaussian, exponential and Poisson), we wrote both a naive Python and an optimized CPython version of their Kullback-Leibler divergence (\ie, $\KL(x,y)$) and a bisection search for the optimization problem for the \klUCB{} index (\ie, $\sup_q \KL(x,q) \leq z$).

\textbf{Variants on $\UCB_1$.}
%
Many variants of the $\UCB_1$ algorithms have been studied in the multi-armed bandits literature.
For instance, \UCB-H replaces the $\log(t)$ by a $\log(T)$ (as well as \klUCB-H), and it actually corresponds to the first algorithm analyzed in \cite{Auer02}.
Using an estimator for the variance of the arms' samples gives \UCB-V, which was proven to be efficient against Gaussian distributions with unknown variance.
Other variants include \UCB$^+$, $\UCB^{\dagger}$ from \cite{Lattimore2018refining}, {\UCB}oost, and many more.
%
All these variants are usually comparable to the original $\UCB_1$ algorithm, in terms of time and memory complexity, and they achieve improved regret upper bounds in the same setting or some extensions of the initial setting.

@@

\begin{table}[!h]
    % \begin{small}  % WARNING
        \centering
        \begin{tabular}{c|cc|cc}
        \textbf{Algorithms}
        & Rewards
        & Regret for $T=100$
        & Rewards
        & Regret for $T=10000$ \\
        % & $\sum\limits_{t=1}^{T=100} r(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r(t)$ & $R_{T=10000}^{\cA}$ \\
        % & $\sum\limits_{t=1}^{T=100} r(t)$ & $R_{T=100}^{\cA}$ & $\sum\limits_{t=1}^{T=10000} r(t)$ & $R_{T=10000}^{\cA}$ \\
            \hline
            Pure exploration
                & $46$ & $20.9$
                & $5171$ & $1886.4$ \\
            \textbf{Pure exploitation}
                & $\mathbf{65}$ & $\mathbf{1.15}$
                & $\mathbf{6997}$ & $\mathbf{0.95}$ \\
            \hline
            $\varepsilon$-greedy
                & $50$ & $20.9$
                & $6657$ & $382.2$ \\
            Explore-then-Commit
                & $49$ & $15$
                & $5766$ & $1199.8$ \\
            \hline
            $\UCB_1$
                & $63$ & $10.3$
                & $6920$ & $78.4$ \\
            \klUCB{}
                & $56$ & $12.4$
                & $6922$ & $70.2$ \\
            \hline
            Thompson sampling
                & $51$ & $12.2$
                & $6949$ & $44.2$ \\
        \end{tabular}
        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{for only one run of the simulation}, for different algorithms.}
        \label{table:2:oneShotResult}
    % \end{small}  % WARNING
\end{table}

@@

\begin{table}[ht]
    % \begin{small}  % WARNING
        \centering
        \begin{tabular}{c|cc|cc}
        \textbf{Algorithms}
        & Rewards
        & Regret for $T=100$
        & Rewards
        & Regret for $T=10000$ \\
        % & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=100} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=100}$ & $\frac{1}{N} \sum\limits_{j=1}^N \sum\limits_{t=1}^{T=10000} r^{(j)}_{A(t)}(t)$ & $\frac{1}{N} \sum\limits_{j=1}^N R^{\cA,(j)}_{T=10000}$ \\
            \hline
            Pure exploration
                & $51$ & $19$
                & $5099$ & $1900$ \\
            Pure exploitation
                & $62$ & $8.3$
                & $6349$ & $653$ \\
            \hline
            $\varepsilon$-greedy
                & $51$ & $18.7$
                & $6460$ & $542$ \\
            Explore-then-Commit
                & $57$ & $12.9$
                & $5795$ & $1207$ \\
            \hline
            $\UCB_1$
                & $\mathbf{60}$ & $\mathbf{9.8}$
                & $6921$ & $79$ \\
            \klUCB{}
                & $61$ & $9.3$
                & $6927$ & $72$ \\
            \hline
            Thompson sampling
                & $\mathbf{60}$ & $\mathbf{9.6}$
                & $\mathbf{6951}$ & $\mathbf{49}$ \\
        \end{tabular}
        \caption{Cumulated rewards and regret, for horizons $T=100$ and $T=10000$, \emph{averaged over $N=1000$ independent simulations} ($j=1,\dots,N$), for different algorithms.}
        \label{table:2:meanResults}
    % \end{small}  % WARNING
\end{table}

@@

\begin{figure}[h!]  % [htbp]
	% \centering
    \includegraphics[width=1.23\linewidth]{SP__K5_T10000_N1000__7_algos/main_MeanRewards____env1-1_2506036032481767447.pdf}
	\caption{Average of the cumulated rewards, as function of $t$, for $T=10000$ and $N=1000$.}
	\label{fig:2:meanRewardsAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
\end{figure}

\begin{figure}[h!]  % [htbp]
	% \centering
	\includegraphics[width=1.17\linewidth]{SP__K5_T10000_N1000__7_algos/main____env1-1_2506036032481767447.pdf}
	\caption{Mean regret $R_t$ as function of $t$ for $T=10000$ and $N=1000$.}
	\label{fig:2:meanRegretAsFunctionOfTimeForDifferentAlgorithmsT10000N1000}
\end{figure}

@@

We presented the simplest MAB model studied in this chapter, by focussing on one agent playing a stationary game.
Both hypotheses are removed or weaken in the rest of this manuscript,
% first in Chapter~\ref{chapter:4} where we consider many XXX
first by considering players facing a stationary problem in a decentralized way in Chapter~\ref{chapter:5},
and then by considering a single player facing a non-stationary or a piece-wise stationary problem in Chapter~\ref{chapter:6}.
%
For both directions, we present in the two final chapters of this thesis natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results for the two problems
of stationary multi-players and piece-wise stationary bandits.

@@

\paragraph{Possible future works.}
%
We focus in this thesis on finite-arms and one-dimensional bandit problems,
and therefore, two possible directions of future works could be to extend our works
to MAB models with either multidimensional rewards, like contextual bandits \cite{Li10,Luo18,ChenLeeLuoWei2019}, or infinite arms \cite{valko2016bandits,Combes17}, like Lipschitz bandits.

### 2019-07-03 2-Chapters/3-Chapter/chapter3.tex
Lots of tiny typos on appendix
Done for 2day

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.75\linewidth]{exampleOfRewards.pdf}
	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$, $0.5$ and $0.9$.}
	\label{fig:3:exampleOfRewards}
\end{figure}

@@

% https://tex.stackexchange.com/a/12430/
\begin{small}
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}
configuration["policies"] = [
  {"archtype": klUCB, "params": {"klucb": klucbBern}},
  {"archtype": klUCBPlusPlus,
   "params": {"horizon": T, "klucb": klucbBern}},
  {"archtype": UCBalpha, "params": {"alpha": 1}},
  {"archtype": Thompson, "params": {"posterior": Beta}}
]
    \end{minted}
    \caption{Example of Python code to configure the list of algorithms tested on a problem.}
    % \captionof{listing}{Small snippet of Python code to configure the list of algorithms tested on a problem. \label{lst:3:howToConfigureAlgorithms}.}
    \label{lst:3:howToConfigureAlgorithms}
\end{listing}
\end{small}

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=1.05\linewidth]{3.pdf}
	\caption[Example of a single-player simulation showing the average regret of $4$ algorithms.]{
		Example of a single-player simulation showing the average regret of four algorithms. They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proven to match the lower-bound up-to a constant), and each but UCB is known to be asymptotically optimal (\ie, with the constant matching the lower-bound).
	}
	\label{fig:3:firstPlot}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=1.05\linewidth]{3_hist.pdf}
	\caption{Histogram of regret for the same experiment as of Figure~\ref{fig:3:firstPlot}. For instance, Thomson sampling is very efficient in average (in yellow), and UCB shows a larger variance (in red).}
	\label{fig:3:firstPlot_hist}
\end{figure}

@@

\subsection{List of research works using SMPyBandits}

@@

We focused here on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
They are more recent (\eg, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or based on a different point-of-view.
The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with or more efficient than other recent algorithms, but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).

@@

\subsection{Computational time.}

@@

\subsection{Memory cost.}

>>>>

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.75\linewidth]{exampleOfRewards.pdf}
	\caption{Histogram of $10000$ \iid{} rewards obtained from three arms with a Gaussian distribution truncated to $[0,1]$, of respective means $0.1$ (in \textcolor{red}{red}), $0.5$ (in \textcolor{green}{green}) and $0.9$ (in \textcolor{blue}{blue}).}
	\label{fig:3:exampleOfRewards}
\end{figure}

@@

% https://tex.stackexchange.com/a/12430/
\begin{small}
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}
configuration["policies"] = [
    { "archtype": klUCB, "params": { "klucb": klucbBern } },
    { "archtype": klUCBPlusPlus, "params": {"horizon": T, "klucb": klucbBern}},
    { "archtype": UCBalpha, "params": { "alpha": 1 } },
    { "archtype": Thompson, "params": { "posterior": Beta } }
]
    \end{minted}
    \caption{Example of Python code to configure the list of algorithms tested on a problem.}
    % \captionof{listing}{Small snippet of Python code to configure the list of algorithms tested on a problem. \label{lst:3:howToConfigureAlgorithms}.}
    \label{lst:3:howToConfigureAlgorithms}
\end{listing}
\end{small}

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=1.05\linewidth]{3.pdf}
	\caption[Example of a single-player simulation showing the average regret of $4$ algorithms.]{
		Example of a single-player simulation showing the average regret of four algorithms (\textcolor{red}{\UCB}, \textcolor{blue}{\KLUCBpp}, \textcolor{darkgreen}{\klUCB} and \textcolor{gold}{Thompson sampling}). They all perform very well: each algorithm is known to be order-optimal (\ie, its regret is proven to match the lower-bound up-to a constant), and each but UCB is known to be asymptotically optimal (\ie, with the constant matching the lower-bound).
	}
	\label{fig:3:firstPlot}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=1.05\linewidth]{3_hist.pdf}
	\caption{Histogram of regret for the same experiment as of Figure~\ref{fig:3:firstPlot}. For instance, \textcolor{gold}{Thomson sampling} is very efficient in average (\textcolor{gold}{in yellow}), and \textcolor{red}{\UCB} shows a larger variance (\textcolor{red}{in red}).}
	\label{fig:3:firstPlot_hist}
\end{figure}

@@

\subsection{List of research works using SMPyBandits}
\label{sub:3:listResearchWorksUsingSMPyBandits}

@@

We focused here on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
They are more recent (\eg, $3$ from $2018$ and $2$ from $2019$) and are usually more complex, or based on a different point-of-view.
The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with or more efficient than other recent algorithms, but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).

@@

\subsection{Computational time}

@@

\subsection{Memory cost}

### 2019-07-03 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done for 2day

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.70\linewidth]{system_model1.eps}
    \caption[In our system model, some dynamic devices (in the IoT network) transmit packets to a gateway and suffer from the interference generated by neighboring networks.]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange} left/right).}
    \label{fig:41:system_model1}
\end{figure}

@@

\begin{figure}[!h]
    \centering
    \subfloat[][10\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{10intelligent.eps}
    \label{fig:41:10intelligent}}
    %
    \subfloat[][30\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{30intelligent.eps}
    \label{fig:41:30intelligent}}
    %
    \vspace*{-10pt}
    \subfloat[][50\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{50intelligent.eps}
    \label{fig:41:50intelligent}}
    %
    \subfloat[][100\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{100intelligent.eps}
    \label{fig:41:100intelligent}}
    \caption{Performance of two MAB algorithms (\UCB{} and Thompson Sampling), compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
    \label{fig:41:from10to100}
    \vspace*{-10pt}
\end{figure}

@@

We can see in Figure~\ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue}), when the number of end-devices is below 50\%. When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
Moreover, we can see in Figures~\ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and suboptimal oracle policies.
For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
Increasing the number of end-devices decreases the gap between the optimal and random policies:
as expected intuitively, the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and on the right end side of Figure~\ref{fig:41:perf_learning}.

\begin{figure}[!h]
    \centering
    \includegraphics[scale=0.80]{perf_learning.eps}
    \caption{Learning with \UCB{} and Thomson Sampling, with many dynamic devices.
        The learning gain, for each device, decreases with the proportion of dynamic devices in the network.
        Note that the values ($16\%$ etc) depend on the repartition of static devices into the $K$ channels (\ie, $S_1,\dots,S_K$) but the general profile of this plot does \emph{not} depend much on these parameters.
        % Both algorithms achieve close to optimal performance after a reasonable learning time.
    }
    \label{fig:41:perf_learning}
\end{figure}

@@

\begin{figure}[!h]
    \centering
    \subfloat[][10\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_10.eps}
    \label{fig:41:ch2_10}}
    %
    \subfloat[][30\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_30.eps}
    \label{fig:41:ch2_30}}
    %
    \vspace*{-10pt}
    \subfloat[][50\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_50.eps}
    \label{fig:41:ch2_50}}
    %
    \subfloat[][100\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_100.eps}
    \label{fig:41:ch2_100}}
    \caption{Performance of the UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.}
    \label{fig:41:figure4appendix}
\end{figure}

>>>>

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.70\linewidth]{system_model1.eps}
    \caption[In our system model, some dynamic devices (in the IoT network) transmit packets to a gateway and suffer from the interference generated by neighboring networks.]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange left/right}).}
    \label{fig:41:system_model1}
\end{figure}

@@

\begin{figure}[!h]
    \centering
    \subfloat[][10\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{10intelligent.eps}
    \label{fig:41:10intelligent}}
    %
    \subfloat[][30\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{30intelligent.eps}
    \label{fig:41:30intelligent}}
    %
    \vspace*{-10pt}
    \subfloat[][50\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{50intelligent.eps}
    \label{fig:41:50intelligent}}
    %
    \subfloat[][100\% of dynamic devices]{\includegraphics[width=0.52\textwidth]{100intelligent.eps}
    \label{fig:41:100intelligent}}
    \caption[Performance of two MAB algorithms, \UCB{} and Thompson Sampling in red, compared to extreme reference policies without learning or oracle knowledge, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.]{Performance of two MAB algorithms, \textcolor{blue}{\UCB{} in blue} and \textcolor{red}{Thompson Sampling in red}, compared to extreme reference policies without learning or \textcolor{gold}{oracle knowledge}, when the proportion of dynamic end-devices in the network increases, from $10\%$ to $100\%$.}
    \label{fig:41:from10to100}
    \vspace*{-10pt}
\end{figure}

@@

We can see in Figure~\ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red $+$}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue} $*$), when the number of end-devices is below 50\%. When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
Moreover, we can see in Figures~\ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and suboptimal oracle policies.
For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
Increasing the number of end-devices decreases the gap between the optimal and random policies:
as expected intuitively, the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and on the right end side of Figure~\ref{fig:41:perf_learning}.

\begin{figure}[!h]
    \centering
    \includegraphics[scale=0.80]{perf_learning.eps}
    \caption[Learning with \UCB{} and Thomson Sampling, with many dynamic devices]{Learning with \textcolor{red}{\UCB} and \textcolor{gold}{Thomson Sampling}, with many dynamic devices. The learning gain, for each device, decreases with the proportion of dynamic devices in the network. Note that the values ($16\%$ etc) depend on the repartition of static devices into the $K$ channels (\ie, $S_1,\dots,S_K$) but the general profile of this plot does \emph{not} depend much on these parameters.}
    \label{fig:41:perf_learning}
\end{figure}

@@

\begin{figure}[!h]
    \centering
    \subfloat[][10\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_10.eps}
    \label{fig:41:ch2_10}}
    %
    \subfloat[][30\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_30.eps}
    \label{fig:41:ch2_30}}
    %
    \vspace*{-10pt}
    \subfloat[][50\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_50.eps}
    \label{fig:41:ch2_50}}
    %
    \subfloat[][100\% of intelligent devices]{\includegraphics[width=0.47\textwidth]{ch2_100.eps}
    \label{fig:41:ch2_100}}
    \caption[Performance of the \UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.]{Performance of the \textcolor{blue}{\UCB} bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.}
    \label{fig:41:figure4appendix}
\end{figure}

### 2019-07-03 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Done for 2day

\begin{figure}[!h]
	\centering
    \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) are sufficient for the two learning devices to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) device, which stays around $40\%$ of success. Similar gains of performance were obtained in other scenarios.}
    \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
\end{figure}

>>>>

\begin{figure}[!h]
	\centering
    \includegraphics[height=9.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
    \caption{Less than $400$ communication slots (\ie, less than $100$ trials in each channel) are sufficient for the two learning devices (\textcolor{red}{\UCB} and \textcolor{green}{Thompson Sampling}) to reach a successful communications rate close to $80\%$, which is \textbf{twice as much} as the \textcolor{cyan}{non-learning (uniform)} device, which stays around $40\%$ of success. Similar gains of performance were obtained in other scenarios.}
    \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
\end{figure}
